Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-05-31 SIR_manuscript.tex
addressing comments, edit figs

\newcommand{\fref}[1]{Fig.~\ref{#1}}
\newcommand{\Rlogo}{R\xspace}
\newcommand{\percap}{\emph{per capita}\xspace}
\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
\newcommand{\covid}{COVID-19\xspace}
\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
\newcommand\pder[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder[x]{y}

@@

The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We present $\Rnum$ in the following form.
\begin{equation}
\label{R0}
\Rnum= \frac{\beta}{\gamma} (1-\Delta), 
\end{equation}
where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation \citep{keeling2011modeling}, and the positive quantity $\Delta$ is the effectiveness of control due to testing and isolation processes defined as 
\begin{equation}
  \label{eq:del4a}
  \Delta= \frac{1}{C N_0}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
\end{equation}
\bmb{super-fussy, but: I don't think we should introduce explicit thinspaces everywhere; why not let
  TeX do the typesetting itself? Do we really prefer more compact equations, and if so is there a package
  that will just change the appropriate settings globally?}\ali{are we talking about \,? not sure about the package!}
where
\begin{align}
\label{eq:C12}
C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \hat F_I,\\
C_2 &= \Big( \omega\gamma(1+p_I)\hat F_I+\gamma^2(\omega+\gamma+\hat F_I)\Big)\theta\_w + \omega^2p_I\hat F_I\theta\_c.
\end{align}
Further details of derivation of $\Rnum$ are provided in appendix, Sec.\,\ref{app:R0}. 
The explicit formula of $\Rnum$ enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$. In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (Sec.\,\ref{app:rho}). 
We derived general expressions for these derivatives. However, we analyzed the effect of $\omega$ on $\Rnum$ for the special case of low testing intensity. Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Rnum$ and $\omega$.

@@

The low-testing case \fref{pan} reflects the case where $rho$ is small relative to the population size. Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$. This testing intensity reflects realistic testing rates during the \covid pandemic, i.e. a maximum of 1.3\% of the population per day, giving a maximum of 10000 tests/day in a population of size $N_0=10^6$ (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021). The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for some population undergoing focused testing, such as a sports league or university.
  The critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is shown as a dotted line in \fref{pan} and \fref{pan2}; combinations of $\rho$ and $\omega$ that lie above this critical contour will control the epidemic. The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is $\approx 1.017$, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing. \bmb{how did we choose this particular value?}\ali{I was thinking of ``critical'' $\Rnum=1$ and wanted to have the baseline in the plot.}
  \bmb{JD wonders whether we really need the ``critical'' $\Rnum=1$ contour or not; any contour will do to illustrate a threshold level of testing intensity, and the fact that $\Rnum=1$ falls in the middle of our range is a bit arbitrary (i.e. the fact that testing can control the epidemic depends on the baseline $\Rnum$ value we have chosen). (I know it was my idea in the first place, but maybe we could dispense with it?)}\ali{agree with the point, the plots will be modified.}

@@

\bmb{can we connect each of these results to specific equations in the appendix?}
\begin{enumerate}

\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always decreases \Rnum;
\item \label{p1:rho} Higher testing intensity $\rho$ decreases $\Rnum$ if
%% \begin{enumerate} \item
testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small.
%% \end{enumerate}
\item \label{p1:omega} Increasing the rate of test return ($\omega$) always decreases \Rnum if waiting individuals do not isolate ($\theta_w=0$).
\item \label{p1:w} Increasing testing focus, i.e. changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always decreases \Rnum.
\end{enumerate}

\newpage
\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
\caption{}\label{p.a}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
\caption{}\label{p.b}
\end{subfigure}
\caption{
{\bf A comparison of effectiveness control on $\Rnum$ at low level of \percap testing intensity $\rho$.}
Numerical evaluation of the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to changes in testing and isolation. Parameter values (Table~\ref{tab:params}):
$N_0=10^6$, $\beta=0.339$ day$^{-1}$ $\gamma= 1/3$ day$^{-1}$; $\omega \in [1/12,2]$ day$^{-1}$;  $\rho \in [0,0.013]$ day$^{-1}$; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$. Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$). Black dotted lines in each panel represent the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ such that $\Rnum=1$.
\bmb{Perhaps we can also include baseline $\Rnum$ and $r$ (or doubling time), i.e. $\beta/\gamma$ and $\beta-\gamma$. (Why did we pick these particular beta/gamma values?)
If $\rho$ is per capita then does the figure depend on $N_0$ at all/can we leave it out of the list of specified parameters?}
}
\label{pan}
\end{figure}

\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_random2.pdf}
\caption{}
\end{subfigure}
%
\begin{subfigure}[t]{.45\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_TTI2.pdf}
\caption{}
\end{subfigure}
\caption{
{\bf A comparison of effectiveness control on $\Rnum$ at high level of \percap testing intensity $\rho$.}
We numerically evaluate the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to testing and isolation. We use the following parameters (see Table~\ref{tab:params} for a brief description):
$N_0=1 \times 10^6$, $\omega \in [1/5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$. Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$. The black dotted line in each panel represents the critical contour of $\Delta=1-\frac{\gamma}{\beta}$, i.e., the $\Delta$ corresponding to the threshold of $\Rnum=1$. \bmb{consider removing dotted line (see note elsewhere); on this and prev figure consider removing redundant y-axis label on right panel (and adjusting/squashing interpanel spacing)?}}
\label{pan2}
\end{figure}

@@

\item Lastly, we show that increasing the focus of testing from random to targeted, always increases $\Delta$, thus decreases \Rnum. This can be seen in \fref{pan} and \fref{pan2}. In particular, we realize higher $\Delta$ levels, i.e., darker colors in corresponding plots\/facets as we go from panel~(a) to panel~(b) in the figures. \bmb{this is the same as point 4 above, right? Am I missing something?}
\end{itemize}

@@

{\bf The first goal} is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$. Mathematically we would like to verify the sign of $\pder[\Rnum]{\theta\_w}$ and $\pder[\Rnum]{\theta\_c}$. We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.\,\eqref{eq:del4}. Thus, $\Delta$ can be rewritten as
\begin{equation}
\label{eq:del4_theta}
\Delta= \frac{1}{C}\Big(
-D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}+
(\omega+\gamma\,S\_u^*)\omega\,p_I\,\hat F_I\,\boldsymbol{\theta\_c}
\Big),
\end{equation}
where
\begin{align}
\label{eq:del4_d1d2}
D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega\,p_I)\gamma\,\hat F_I, \\
D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega\,p_I)\gamma\,\hat F_I+(\gamma+p_I\,\hat F_I)\omega^2.
\end{align}
$\Delta$, Eq.\,\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient given by 
$$1/C\,(\gamma\,S\_u^*\,+\omega(1-\theta\_w\,S\_u^*))\omega\,p_I\,\hat F_I.$$
This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder[\Rnum]{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.

With a similar logic, $\Delta$ \eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
\begin{equation}
\label{eq:del4_thetaw}
1/C\,\Big( -D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w} \Big).
\end{equation}
We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder[\Rnum]{\theta\_w}\leq 0$.
It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
\begin{align}
\label{eq:d_del4_thetaw}
\pder[\Delta]{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
1/C\Big( (D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \Big) \\
=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2\,p_I\,\hat F_I )\,S\_n^*
+\gamma(\omega+\gamma)\,\hat F_I\,(1-S\_n^*) \Big),
\end{align}
which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.

{\bf The second goal} is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$. Mathematically we would like to verify the sign of $\pder[\Rnum]{\rho}$, which specifically depends on $\pder[\Delta]{\rho}$. We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.\,\eqref{dfe}, in $\Delta$ \eqref{eq:del4}. Also, we define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$. This is mainly to avoid singularity in $\hat F_I$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$. Thus, $\rho$ is reparameterized as 
\begin{equation}
\label{eq:phi}
\rho=\frac{\omega \phi}{\omega+\phi}.
\end{equation}
This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder[\Delta]{\phi}$ instead of the complicated $\pder[\Delta]{\rho}$.
The derivative is
\begin{align}
\label{eq:dd3dr}
\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3\,\phi^2+b_3\,\phi+c_3),
\end{align}
where
\begin{align}
\label{eq:abcd2}
a_3& = \frac{w_I}{w_S} \Bigg(&& (1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w\,\gamma^3
+(1-\theta\_c)\,p_I^2\,\theta\_w\,\frac{w_I}{w_S}\,\omega^3 \\ \notag
& ~ &&+\Big( \big( (1-\theta\_w-\frac{w_I}{w_S})\,\theta\_c +(3-2\theta\_w)\,\theta\_w\frac{w_I}{w_S} \big)\,p_I
+(1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w  \Big)\,\omega\,\gamma^2 \\ \notag
& ~ &&+\Big(
\big((1-\theta\_w-\theta\_w\,\frac{w_I}{w_S})\,\theta\_c+ (2-\theta\_w)\,\theta\_w\,\frac{w_I}{w_S} \big)\,p_I
+(2\,\theta\_w-\theta\_w^2-\theta\_c)\,\frac{w_I}{w_S}\,p_I^2
\Big)\,\omega^2\,\gamma \Bigg), \\
b_3& = && \hspace{-1cm} 2\frac{w_I}{w_S}\,(\omega+\gamma)\gamma\Big(
(\omega+\gamma+\omega\,p_I)(2-\theta\_w)\,\gamma\,\theta\_w+ (1-\theta\_w)\,\omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
\Big) ,\\
c_3& = && \hspace{-1cm} (\omega+\gamma)^2\,\gamma\Big(
(2-\theta\_w)\,\gamma^2\,\theta\_w+
(1+\frac{w_I}{w_S})\,\omega\,\gamma\,\theta\_w+
\frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}\, \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
\end{align}
Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$. It is straight forward to show that $a_3\geq 0$ in case of random testing strategy, i.e., $w\_S=w\_I=1$. 
If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.

>>>>

\newcommand{\fref}[1]{Fig.~\ref{#1}}
\newcommand{\Rlogo}{R\xspace}
\newcommand{\percap}{\emph{per capita}\xspace}
\newcommand{\Rnum}{\ensuremath{\mathcal{R}_0}\xspace}
\newcommand{\covid}{COVID-19\xspace}
\newcommand{\pro}[1]{\ensuremath{\frac{\partial #1}{\partial \rho}}}
\newcommand\pder[2]{\ensuremath{\frac{\partial#1}{\partial#2}}} %\pder y{x}

@@

The basic reproduction number, $\Rnum$, was calculated by using the next-generation matrix method \citep{van2002reproduction}. We present $\Rnum$ in the following form.
\begin{equation}
\label{R0}
\Rnum= \frac{\beta}{\gamma} (1-\Delta), 
\end{equation}
where the term $\frac{\beta}{\gamma}$ is the classical basic reproduction number for a SIR model without testing and isolation \citep{keeling2011modeling}, and the positive quantity $\Delta$ is the effectiveness of control due to testing and isolation processes defined as 
\begin{equation}
  \label{eq:del4a}
  \Delta= \frac{1}{C N_0}\big(C_1 S\_u^*+(C_2(1-\theta\_w)+C \theta\_w)S\_n^*\big),
\end{equation}
\bmb{super-fussy, but: I don't think we should introduce explicit thinspaces everywhere; why not let
  TeX do the typesetting itself? Do we really prefer more compact equations, and if so is there a package
  that will just change the appropriate settings globally?}\ali{are we talking about \\,?}
where
\begin{align}
\label{eq:C12}
C_{\phantom 1} &= (\omega+\gamma) \Big(\gamma(\omega+\gamma)+(\gamma+\omega p_I)\hat F_I\Big),\\
C_1 &= (\omega+\gamma)(\theta\_w \gamma+\theta\_c \omega p_I) \hat F_I,\\
C_2 &= \Big( \omega\gamma(1+p_I)\hat F_I+\gamma^2(\omega+\gamma+\hat F_I)\Big)\theta\_w + \omega^2p_I\hat F_I\theta\_c.
\end{align}
Further details of derivation of $\Rnum$ are provided in appendix, Sec.\,\ref{app:R0}. 
The explicit formula of $\Rnum$ enables us to study the effects of testing and isolation parameters on $\Rnum$ both analytically and via numerical solutions.
We are specifically interested in parameters that could be manipulated by public health policy: isolation efficacy, $\theta\_c$ and $\theta\_w$; \percap testing intensity, $\rho$; and the rate of test return, $\omega$. In particular, we look at the partial derivatives of $\Delta$ with respect to these parameters (Sec.\,\ref{app:rho}). 
We derived general expressions for these derivatives. However, we analyzed the effect of $\omega$ on $\Rnum$ for the special case of low testing intensity. Specifically, by making the restriction $\rho \ll 1$, we are able to Taylor-expand $\Delta$ at $\rho=0$ and analyze the resulting simplified derivatives to illustrate a surprising non-monotonic relationship between $\Rnum$ and $\omega$.

@@

The low-testing case \fref{pan} reflects the case where $rho$ is small relative to the population size. Specifically, $\rho \in [0,0.013]$, and test return rate $\omega\in [1/12,2]$. This testing intensity reflects realistic testing rates during the \covid pandemic, i.e. a maximum of 1.3\% of the population per day, giving a maximum of 10000 tests/day in a population of size $N_0=10^6$ (this is approximately four times the maximum testing rate in Ontario, Canada in mid-2021). The less realistic high-testing case \fref{pan2} is included to highlight the occurrence of non-monotonic changes in $\Rnum$ with respect to $\rho$.
In \fref{pan2} the maximum capacity of $\rho$ is larger relative to the population size, $\rho \in [0,1/5)$ and the test return rate $\omega\in [1/5,2]$; these values are clearly unrealistic for a large population but might be relevant for some population undergoing focused testing, such as a sports league or university.
  The critical contour of $\Delta=1-\frac{\gamma}{\beta}$, corresponding to the threshold $\Rnum=1$, is shown as a dotted line in \fref{pan} and \fref{pan2}; combinations of $\rho$ and $\omega$ that lie above this critical contour will control the epidemic. The implied baseline value of $\Rnum=\frac{\beta}{\gamma}$ in these figures is $\approx 1.017$, corresponding to a case where the epidemic spreads slowly even in the absence of testing and tracing. \bmb{how did we choose this particular value?}\ali{I was thinking of ``critical'' $\Rnum=1$ and wanted to have the baseline in the plot.}
  \bmb{JD wonders whether we really need the ``critical'' $\Rnum=1$ contour or not; any contour will do to illustrate a threshold level of testing intensity, and the fact that $\Rnum=1$ falls in the middle of our range is a bit arbitrary (i.e. the fact that testing can control the epidemic depends on the baseline $\Rnum$ value we have chosen). (I know it was my idea in the first place, but maybe we could dispense with it?)}\ali{agree, the plots are updated.}

@@

\item \label{p1:eta} Increasing isolation efficacy for waiting ($\theta\_w$) and confirmed-positive ($\theta\_c$) individuals always decreases \Rnum (Eqs.~\ref{eq:del4_theta},~\ref{eq:d_del4_thetac},~\ref{eq:d_del4_thetaw});
\item \label{p1:rho} Higher testing intensity $\rho$ decreases $\Rnum$ if
%% \begin{enumerate} \item
testing is random (all $w_X$ equal) or testing intensity ($\rho$) is small (Eq.~\ref{eq:dd3dr}).
%% \end{enumerate}
\item \label{p1:omega} Increasing the rate of test return ($\omega$) always decreases \Rnum if waiting individuals do not isolate ($\theta_w=0$) (Eq.~\ref{eq:dlindo}).
\item \label{p1:w} Increasing testing focus, i.e. changing the testing weights from random ($w_S=w_I$) toward targeted  ($w_S<w_I$), always decreases \Rnum (Eq.~\ref{eq:d_del_wis}).
\end{enumerate}

\newpage
\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.46\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_random.pdf}
\caption{}\label{p.a}
\end{subfigure}
%
\begin{subfigure}[t]{.44\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_TTI.pdf}
\caption{}\label{p.b}
\end{subfigure}
\caption{
{\bf A comparison of effectiveness control on $\Rnum$ at low level of \percap testing intensity $\rho$.}
Numerical evaluation of the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to changes in testing and isolation. Parameter values (Table~\ref{tab:params}):
$\beta=0.339$ day$^{-1}$ $\gamma= 1/3$ day$^{-1}$; $\omega \in [1/12,2]$ day$^{-1}$;  $\rho \in [0,0.013]$ day$^{-1}$; $\theta\_w$ and $\theta\_c$ vary between 0 (no effect of isolation) and 1 (complete elimination of transmission); $p_S=0$, $p_I=1$ and $p_R=0.5$. Contours of $\Delta$ are plotted for (a) random testing ($w_S=w_I=w_R=1$) and (b) targeted testing ($w_S=0.3$; $w_I=w_R=1$). 
The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e. $\beta/\gamma$ and $\beta-\gamma$ respectively, are approximately 1.017 and 0.006 in this simulation.
}
\label{pan}
\end{figure}

\begin{figure}[h!]
\centering
\begin{subfigure}[t]{.46\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_random2.pdf}
\caption{}
\end{subfigure}
%
\begin{subfigure}[t]{.44\textwidth}
\centering
\includegraphics[width=\linewidth]{codes/R0contour_TTI2.pdf}
\caption{}
\end{subfigure}
\caption{
{\bf A comparison of effectiveness control on $\Rnum$ at high level of \percap testing intensity $\rho$.}
We numerically evaluate the effectiveness control parameter, $\Delta$ (Eq.~\ref{eq:del4}), with respect to testing and isolation. We use the following parameters (see Table~\ref{tab:params} for a brief description):
$\omega \in [1/5,2]$ 1/day, $1/\gamma= 3$ days, $\rho \in [0,1/5)$ 1/day, $\theta\_w$ and $\theta\_c$ vary between 0 and 1 with 0 for no effect and 1 for full effect of isolation on the transmission probability, $\beta=0.339$ 1/day, $p_S=0$, $p_I=1$ and $p_R=0.5$. Contours of $\Delta$ are plotted for two testing strategies identified by a set of relative testing weights; (a) random testing where $w_S=w_I=w_R=1$ and (b) targeted testing where $w_S=0.3$ and $w_I=w_R=1$. The baseline basic reproduction number $\Rnum$ and the doubling time $r$, i.e. $\beta/\gamma$ and $\beta-\gamma$ respectively, are approximately 1.017 and 0.006 in this simulation. \bmb{consider removing dotted line (see note elsewhere); on this and prev figure consider removing redundant y-axis label on right panel (and adjusting/squashing interpanel spacing)?}
\ali{ok?}
}
\label{pan2}
\end{figure}

@@

\item Lastly, we show that increasing the focus of testing from random to targeted, always increases $\Delta$, thus decreases \Rnum. This can be seen in \fref{pan} and \fref{pan2}. In particular, we realize higher $\Delta$ levels, i.e., darker colors in corresponding plots\/facets as we go from panel~(a) to panel~(b) in the figures. \bmb{this is the same as point 4 above, right? Am I missing something?}
\ali{you are right.}
\end{itemize}

@@

{\bf The first goal}
is to explore how changes in isolation, $\theta\_w$ and $\theta\_c$, affects $\Rnum$. Mathematically we would like to verify the sign of $\pder \Rnum{\theta\_w}$ and $\pder \Rnum{\theta\_c}$. We start with simplifying $\Delta$ \eqref{eq:del4} by factoring $\theta\_w$ and $\theta\_c$ in Eq.\,\eqref{eq:del4}. Thus, $\Delta$ can be rewritten as
\begin{equation}
\label{eq:del4_theta}
\Delta= \frac{1}{C}\Big(
-D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w}+
(\omega+\gamma\,S\_u^*)\omega\,p_I\,\hat F_I\,\boldsymbol{\theta\_c}
\Big),
\end{equation}
where
\begin{align}
\label{eq:del4_d1d2}
D_1=& (\omega+\gamma)\gamma^2+(\omega+\gamma+\omega\,p_I)\gamma\,\hat F_I, \\
D_2=& (3\omega+2\gamma)\gamma^2+(\omega+\gamma+2\omega\,p_I)\gamma\,\hat F_I+(\gamma+p_I\,\hat F_I)\omega^2.
\end{align}
$\Delta$, Eq.\,\eqref{eq:del4_theta}, is linear in $\theta\_c$ with a positive coefficient. thus
\begin{equation}
\label{eq:d_del4_thetac}
\pder\Delta{\theta\_c}=1/C(\gamma S\_u^*+\omega(1-\theta\_w S\_n^*))\omega p_I \hat F_I.
\end{equation}
This results in increasing $\Delta$, thus decreasing $\Rnum$ with respect to $\theta\_c$, that is $\pder \Rnum{\theta\_c}\leq 0$. Note that $C$ is independent of $\theta\_c$ and $\theta\_w$.

With a similar logic, $\Delta$ \eqref{eq:del4_theta} is a concave-down quadratic equation in $\theta\_w$, given by
\begin{equation}
\label{eq:del4_thetaw}
1/C\,\Big( -D_1\,S\_n^*\,\boldsymbol {\theta\_w^2}+
\big(-\omega^2\,p_I\,\hat F_I\,S\_n^*\,\boldsymbol{\theta\_c}\,+\,D_2\,S\_n^*\,+\,\gamma\,\hat F_I(\omega+\gamma) \big)\boldsymbol{\theta\_w} \Big).
\end{equation}
We show that the feasible range of $\theta\_w$ lies between 0 and the vertex of this parabola where the parabola is increasing in $\theta\_w$, and so does $\Delta$ which results in inferring $\pder\Rnum{\theta\_w}\leq 0$.
It is enough to show that partial derivative of the expression \eqref{eq:del4_thetaw} with respect to $\theta\_w$ at $\theta\_w=1$ is non-negative. It follows that
\begin{align}
\label{eq:d_del4_thetaw}
\pder\Delta{\theta\_w}\bigg\rvert_{\theta\_w=1} =&
1/C\Big( (D_2-2D_1-\omega^2\,p_I\,\hat F_I\,\theta\_c)\,S\_n^*+\gamma\,\hat F_I(\omega+\gamma) \Big) \\\notag
=& 1/C\Big( (\gamma(\omega+\gamma)+\gamma\omega^2+(1-\theta\_c)\omega^2\,p_I\,\hat F_I )\,S\_n^*
+\gamma(\omega+\gamma)\,\hat F_I\,(1-S\_n^*) \Big),
\end{align}
which is a positive quantity, given that $\theta\_c$ and $S\_n^*$ vary between 0 and 1.

{\bf The second goal} is to explore how changes in \percap testing intensity $\rho$ affects $\Rnum$. Mathematically we would like to verify the sign of $\pder\Rnum{\rho}$, which specifically depends on $\pder\Delta{\rho}$. We use the derived expressions for $S\_u^*$ and $S\_n^*$, given by Eqs.\,\eqref{dfe}, in $\Delta$ \eqref{eq:del4}. Also, we define $\phi = \hat F_S = \frac{\rho \omega}{\omega-\rho}$, to reparameterize $\rho$. This is mainly to avoid singularity in $\hat F_I$ \eqref{eq:fi}, when testing intensity $\rho$ is very close to the rate of test return $\omega$. Thus, $\rho$ is reparameterized as 
\begin{equation}
\label{eq:phi}
\rho=\frac{\omega \phi}{\omega+\phi}.
\end{equation}
This one-to-one monotonic reparameterization enables us to simplify the mathematical expressions and explore the simpler $\pder\Delta{\phi}$ instead of the complicated $\pder\Delta{\rho}$.
The derivative is
\begin{align}
\label{eq:dd3dr}
\partial\Delta/\partial\phi=& \frac{1}{d_3} (a_3\,\phi^2+b_3\,\phi+c_3),
\end{align}
where
\begin{align}
\label{eq:abcd2}
a_3& = \frac{w_I}{w_S} \Bigg(&& (1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w\,\gamma^3
+(1-\theta\_c)\,p_I^2\,\theta\_w\,\frac{w_I}{w_S}\,\omega^3 \\ \notag
& ~ &&+\Big( \big( (1-\theta\_w-\frac{w_I}{w_S})\,\theta\_c +(3-2\theta\_w)\,\theta\_w\frac{w_I}{w_S} \big)\,p_I
+(1-\theta\_w)\,(1+\frac{w_I}{w_S})\theta\_w  \Big)\,\omega\,\gamma^2 \\ \notag
& ~ &&+\Big(
\big((1-\theta\_w-\theta\_w\,\frac{w_I}{w_S})\,\theta\_c+ (2-\theta\_w)\,\theta\_w\,\frac{w_I}{w_S} \big)\,p_I
+(2\,\theta\_w-\theta\_w^2-\theta\_c)\,\frac{w_I}{w_S}\,p_I^2
\Big)\,\omega^2\,\gamma \Bigg), \\
b_3& = && \hspace{-1cm} 2\frac{w_I}{w_S}\,(\omega+\gamma)\gamma\Big(
(\omega+\gamma+\omega\,p_I)(2-\theta\_w)\,\gamma\,\theta\_w+ (1-\theta\_w)\,\omega^2 p_I\theta\_c+\omega^2 p_I\theta\_w
\Big) ,\\
c_3& = && \hspace{-1cm} (\omega+\gamma)^2\,\gamma\Big(
(2-\theta\_w)\,\gamma^2\,\theta\_w+
(1+\frac{w_I}{w_S})\,\omega\,\gamma\,\theta\_w+
\frac{w_I}{w_S} \omega^2 p_I\theta\_c \Big),\\
d_3&  = && \hspace{-1cm} \frac{(\omega+\gamma)}{\omega}\, \Big((\omega p_I+\gamma)\frac{w_I}{w_S}\phi\,+\,(\omega+\gamma)\gamma \Big)^2(\omega+\phi)^2.
\end{align}
Note that $\phi\geq 0$, also $b_3$, $c_3$ and $d_3$ are all positive. However $a_3$ can be positive or negative.
If $a_3\geq 0$, $\partial\Delta/\partial\phi \geq 0$ for all feasible range of parameters, thus $\pro\Rnum \leq 0$. It is straight forward to show that $a_3\geq 0$ in case of random testing strategy, i.e., $w\_S=w\_I=1$. 
If $a_3 < 0$, then the quadratic expression in the numerator of \eqref{eq:dd3dr} has a positive root, $\phi^*$, such that for $\phi>\phi^*$, $\partial\Delta/\partial\phi < 0$.

An example of this counter-vailing effect of $\phi$, and consequently $\rho$, on $\Rnum$ occurs when $\theta\_w=0$ and $\theta\_c=1$.
This is illustrated in the top-right panel of the \fref{pan2} panel (b), where the strength of isolation for awaiting people is the least, but the most for the confirmed cases. In this case, simplifying $a_3$ in Eq.~\eqref{eq:abcd2} gives $$a_3=\frac{w_I}{w_S}\,\omega\,\gamma\,p_I((\omega+\gamma)-\frac{w_I}{w_S}(\omega\,p_I+\gamma)).$$
Specifically, in the case of targeted testing which is identified by $\frac{w_I}{w_S}> 1$, and using a perfect sensitive test, thus $p_I=1$, there exists a range for $\rho$ over which $\pder\Rnum{\rho}\leq\,0$.  
Note that $\rho$ and $\omega$ have a similar mechanism in delaying people to get into $I\_c$, thus we would expect to see the non-trivial counter-vailing effect of these two parameters on $\Rnum$.
